---
title: "Wrangling Practice with US President Data"
author: "NAME"
date: "`r format(Sys.Date())`"
output: pdf_document
---


In this assignment, you'll work with the dataset called `presidential` in the `ggplot2` packages.  I'll also introduce you to the `lubridate` packages that is mentioned in your book but will not be covered in class. 


# Setup and Preliminaries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load tidyverse}
library(tidyverse)
```


1. What variables are in the data `presidential`? 
```{r names}
names(presidential)
```

2. How many rows and columns?
```{r dimensions}
dim(presidential)
```

*****
\break

# Practice your Wrangle Skills

3. Use `select` to retrieve only the names and party affiliations for the presidents listed. Try and use the `%>%` operator in all your code as much as you can.  Do not save it as a new object.  Just display.
```{r select name and party}
presidential %>% select(name, party)
```

4. Use `filter` to retrieve the Republican presidents only from the `presidential` database.  Do not save it as a new object.  Just display.
```{r filter party}
presidential %>% filter(party=="Republican")
```

5. What Democratic presidents served since 1973? Only show the names of the presidents.  Use `filter` and `select` with a `%>%`.
```{r filter and select1}
filter(presidential, start > 1973 & party=="Democratic") %>% 
  select(name)
```

5a. Single line for same code with nesting.
```{r filter and select2}
select(filter(presidential, start > 1973 & party == "Democratic"), name)
```

*****
\break

# Working with Dates using `lubridate`

Lubridate works well with our data to manipulate and work with dates. Here you'll use your first `lubridate` functions called `interval` and `dyears`. Here's a <a  href="https://cran.r-project.org/web/packages/lubridate/lubridate.pdf">link</a> to help pages for the library.  


```{r load lubridate}
library(lubridate) #included in tidyverse but not loaded
```

## Using `interval` and `dyears`

6. How long was the president's term? Save it to a new object called `mypresidents`.
```{r term length}
mypresidents <- presidential %>%
  mutate(term.length = interval(start, end) / dyears(1))
mypresidents
```

7. Checkout `?interval` to see what interval does. You can do the same with `?dyears`.
What do they do?
interval:
dyears:

*****
\break

## `mutate` for years elected and retired

8. Practice `rename` to make the existing variable names appear as you want them to appear. In this case, remove the period between the variable we just created called `term.length`.  Rename it to `term_length` and resave it as `mypresidents`.
```{r}
mypresidents <- mypresidents %>% rename(term_length = term.length)
mypresidents
```

9. Create a variable that shows the year the president was elected.  Here we use `year` from `lubridate` within `mutate`.  Just uncomment the commands and run it.
```{r}
mypresidents <- mutate(mypresidents, elected = year(start)-1)
mypresidents
```
What did it do?
Answer:

10. Repeat the above code but create a new variable called `retired` that notes one year after their term ended.
```{r}
mypresidents <- mutate(mypresidents, retired = year(end)+1)
mypresidents
```
What did it do?
Answer:

*****
\break

## Revising unique cells

11. Lyndon Johnson and Gerald Ford were not elected.  We need to sift through the data and put in a `NA` for their elected year.  You can do this using the `ifelse` function and the `%in%` operator. Explain what this doing.
```{r}
mypresidents <- mypresidents %>% 
  mutate(elected = ifelse((name %in% c("Johnson","Ford")), NA, elected))
mypresidents
```

Answer:

12. Sort the data frame by `term_length` using `arrange`. Don't save it just run with with a pipe. Who had the shortest term length?
```{r}
mypresidents %>% arrange(desc(term_length))
```
Answer: 


13. Use `arrange` to sort the further by `party` (still use `term_length`!). Use `arrange`, the pipe, and `select` to show only `name`, `term_length` and `party`.  What democrat had the shortest term?
```{r}
mypresidents %>% 
  arrange(desc(term_length), party) %>% 
  select(name, term_length, party)
```

Answer: 


*****
\break

# Revisiting `summarize` using the presidential data

14. `Summarize` will provide a single row of information that can summarize your data in a number of ways.  In one command using `summarize`, see if you can create the following variables:

* `n()` counts up the number of people
* `first_year` using min(year(start)) to tell us first year.
* `last_year` using max(year(end)) to tell us the last year.
* `avg_term_length` using mean by term_length
* `num_dems` using sum by party=="Democratic" (it adds them up)


## Create a summary of the data
```{r}
mypresidents %>%
  summarize(
    N = n(), 
    first_year = min(year(start)), 
    last_year = max(year(end)),
    avg_term_length = round(mean(term_length), digits=2),
    num_dems = sum(party=="Democratic"), 
    years = round(sum(term_length), digits=2)) 
```
What was the average term length?

*****
\break

## Add `party` as a grouping variable to the above code.

15. `Summarize` the code from above by group it by `party` using `group_by`
```{r}
mypresidents %>%
  group_by(party) %>%
  summarize(
    N = n(), 
    first_year = min(year(start)), 
    last_year = max(year(end)),
    avg_term_length = round(mean(term_length), digits=2),
    num_dems = sum(party=="Democratic"), 
    years = round(sum(term_length), digits=2)) 
```
What was the average term length for Republicans? Democrats?













