---
title: "R Markdown Tips"
author: "Jamie Bedics"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
    toc: true
  pdf_document: default
  html_document:
    toc: true
    toc_float: true
---


## Setup

This chunk pops up for _every_ `.Rmd` you open.  It will controls the options that appear in **every** chunk in your document.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

In this particular setup chunk, it's call the `opts_chunk$set` function from the `knitr` library. I'll never see warnings or messages but I _will_ see code because `echo = True` and the results.  Notice `include=FALSE` will including _nothing_ from the chunk in the output.

The second line also directs you to your R Project as the default working directory.

### 1. Hide code using `echo=FALSE`

a. Here you see code when you render:
```{r code}
a <- 1+1
a
```
Some people might want to see your code and others might not care.


b. If you want to hide your code from the audience then change `echo=` to `FALSE`: 
```{r no code, echo = FALSE}
a <- 1+1
a
```
It's much cleaner this way especially if you have a lot of preliminary work.  


### 2. Common used chunk options

The additional options which are always worth using:

- warning = FALSE
- messages = FALSE


It will clean up a lot of mess in your nicely rendered documents. 


Further chunk options can be found at the following **[site](https://yihui.name/knitr/options)**.

\break

## Page Break

The following are instructions for adding page breaks in your documents.  It's VERY easy for PDF but not so easy with DOC output.

### 1. Add Page Breaks to a rendered .PDF document. 

You can add a page break and create a separate page on pdf if you add the `\break` command.  This does not work for .html or for .docx.

Do not inlcude the `\break' in a chunck and do not include ticks that surround slash and word

### 2. Add Pagre Breaks to a rendered .DOCX document.

If you want to do so for .docx the instructions are pretty straightforward and can be found at this website:
[https://datascienceplus.com/r-markdown-how-to-insert-page-breaks-in-a-ms-word-document/](https://datascienceplus.com/r-markdown-how-to-insert-page-breaks-in-a-ms-word-document/)

The steps can be broken down as follows:

1. Render a word doc and save is as something like "rmarkdown_template.docx" and save in the **same** folder as your code.

2. Open the new word file and change the style template as noted on the website.  There are several steps.

3. Save the template.

4. In the YAML  (the code at the very top with you name), remove the `default` and add the following under word_document: 
  reference_docx: rmarkdown_template.docx

5. Once you complete these steps, you just have to add a five number signs `#####` to indicate a page break.

As an aside, notice the number formmating comes out nicely above but we can use bullets too:

- Render a word doc and save is as something like "rmarkdown_template.docx" and save in the **same** folder as your code.

- Open the new word file and change the style template as noted on the website.  There are several steps.

- Save the template.

- In the YAML  (the code at the very top with you name), remove the `default` and add the following under     
    - word_document: 
        
        reference_docx: rmarkdown_template.docx

- Once you complete these steps, you just have to add a five number signs `#####` to indicate a page break.

\break

## Equations

You can get fancy with equations:

$$\begin{array}{ccc}
x_{11} & x_{12} & x_{13} \\
x_{21} & x_{22} & x_{23}
\end{array}$$

\break

## Quotations

Sometimes in describing your results you might want to quote a famous person for inspiration:

> "Today is only one day in all the days that will ever be. But what will happen in all the other days that ever come can depend on what you do today."
>
> --- Ernest Hemingway, _From Whom the Bell Tolls_

\break

## Table of Contents

Notice in the YAML (top of the page), that under `.html` there is a `toc=true` and there is no `default` statement like there is for word and pdf.  Now, Knit html and .doc and compare. What's the difference?

Add `toc_float = true` under `toc=true` and knit to html.  What happens?  

Additional tips on `html` can be found **[here](https://rmarkdown.rstudio.com/html_document_format)** at R Studio.

#####

\break

## Practice Exercises

1. Create an identical chunk to the one below but hide the code.
```{r practice1}
plot(cars)
```

Hide the code:


2. Make a quote

3. Add a break line so the `Practice Exercises` are on a seperate page if you can render to .pdf. You can do this even if you can't render to .pdf.

4. Add a TOC to the the word document format.

5. Make all the headers one size smaller except for the 5 level header right before `Practical Exercises`



*****



