---
title: "DSUR Chapter 3 Textbook Code"
author: "Jamie Bedics"
date: "9/13/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Creating Variables (3.5.1)

Here we use the concatenate function to create an object.  Here we create a dataframe that consists of the names of members of Metallica's band and their age.

```{r metallica df}
metallicaNames <- c("Lars", "James", "Kirk", "Rob")
metallicaAges <- c(47, 47, 48, 46)
```


### Creating a Dataframe (3.5.2)
```{r}
metallica.df <- data.frame(name = metallicaNames, age = metallicaAges)
metallica.df # .df was added so you know what is a `dataframe`
```

Here we add new information to the _existing_ dataframe. 
 
```{r children}
metallica.df$childAge <- c(12, 12, 4, 6)
```

You will most likely **never** create a dataframe in R unless you're doing a simulation or simply playing with R.


### Calculating a new variable from exisiting (3.5.3)

Create the age of the band member when they had their children within the existing dataframe
```{r}
metallica.df$fatherhoodAge <- metallica.df$age - metallica.df$childAge
```

Working with Dates 
```{r}
metallica.df$anniversary <- as.Date(c("1990-06-01", "2000-12-12", "1994-06-01", "2005-10-12"))
```

Examine the structure of the data.
```{r}
str(metallica.df)
```







