---
title: "R4DS Chapter 1 Textbook Code"
author: "Jamie Bedics"
date: "9/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Prerequisites
```{r}
library(tidyverse) # this assumes I installed the `tidyverse` in the console
```

Tidyverse is a collection of packages that are grouped together.  They include, for example, `ggplot2` which is a library we frequently use for data visualization.  If you have trouble loading the `tidyverse`

If this doesn't work, you could load the packages individually like below:
```{r}
library(ggplot2)
```

## First Steps

### The MPG Dataframe

```{r}
mpg
```

1. How many rows? A:
2. How many Columns? A:
3. What type of data is `manufacturer`? `model`? `disp`? `cty`?


### What's in the data?

Since the data is in a package then there will be a help page associated with it.

```{r}
?mpg
```
1. Describe the dataset.
2. What is `disp` anyway?  


### Creating a ggplot

1. Book example pg. 5
```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x=displ, y=hwy))
```

A different version with x and y in first command. More typical
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point()
```

The first line defines the data and your x and y-axes.  The second line, "geom_point", adds a layer of points.  We can change and add layers. Inside each layer, we learn commands that allow us to adjust the quality of each layer or function.


### Exercises pg. 6
1. 
2.
3.
4.
5.








